Effects of the body force on the pedestrian and the evacuation dynamics

### 2020-02-14 paper.tex
puse en azul cosas que antes estaban en rojo

The numerical \textcolor{blue}{parameters} for the ``faster is slower'' effect presented by 
Helbing and co-workers \textcolor{red}{\sout{is somewhat cumbersome} appears to have some drawbacks} 
Ref.~\cite{helbing_2000,dorso_2017,dorso_2019}. Although a single 
parameter (say, the desired velocity $v_d$) is numerically varied to 
explain the phenomenon, the researcher \textcolor{red}{\sout{looses sight} may loose sight} of the dimensionless 
``numbers'' that truly control the crowd dynamics. The \textcolor{blue}{set of parameters} in 
Ref.~\cite{helbing_2000} also misses the ``faster is faster'' effect reported 
to occur at very high pedestrian densities \cite{dorso_2017,haghani_2019}. 
\textcolor{blue}{Moreover}, the empirical fundamental diagram raises as a point of reference 
for the SFM control parameters \cite{helbing_2007,dorso_2017}. \\

The fundamental diagram exhibits the flux behavior for either low \textcolor{blue}{density} crowds 
(with \textcolor{blue}{rare} contacts between pedestrians) and highly dense crowds (dominated 
by \textcolor{blue}{two body interactions}). \textcolor{blue}{In the second case, crowds} experience a flux 
slowing down, but other behaviors are also possible \cite{helbing_2007,lohner_2018}. \textcolor{red}{
\sout{In light of our previous hypothesis, we } We} may suspect that the  modeling of the 
``flux slowing down'' within the context of the SFM will require the proper 
setting of the (dimensionless) controlling parameters. We \textcolor{blue}{have already}
examined this working hypothesis in Ref.~\cite{dorso_2019}, but limiting the parameter exploration to 
the sliding friction, disregarding the body force. 
We now widen the investigation on the \textcolor{blue}{\sout{parameter settings} set of parameters} to include the one 
associated to the body force. We will  explore 
 the complex interplay between 
the body force and the sliding friction among pedestrians. Recall that the 
interplay dynamics is not directly controlled by the \textcolor{blue}{\sout{numerical setting} set of parameters}, but 
through dimensionless ``numbers'', where the model parameters appear mixed 
between each other. Thus, this step up offers a challenge to the 
``individualistic'' meaning of the parameter's \textcolor{blue}{\sout{setting} set}. \\

@@

The complex behavior of pedestrians features
 either his (her) feelings and 
the environmental conditions. The former is expressed, for example, by his 
(her) moving ``attitude'' (say,\textcolor{red}{\sout{self-assuredness} relaxed, normal, aggressive, etc}).
The latter brings out the 
observed separation between pedestrians. Also, the ``contacts'' between 
individuals \textcolor{blue}{produce} some kind of ``unwanted'' slowing down. All these 
observed patterns are commonly quantified in the literature into a set of 
characteristic  parameters. \textcolor{red}{It is worth mentioning that the empirical
data reported in this Section corresponds to non-panicking situations in which pedestrians 
do not experience high anxiety levels (see below). }
The experimental meaning of these parameters is as follows.

\begin{itemize}
\item[(i)] \textcolor{red}{\textbf{The walking attitude of a pedestrian may appear somewhat 
  ``aggressive'' if he (she) reacts actively to unexpected behaviors 
\cite{lakoba_2005,helbing_1995}. The smaller the reaction time, the more 
 aggressive observed posture. The associated parameter to this 
behavior is the relaxation or characteristic time $\tau$ 
\cite{johansson_2009,helbing_2000}}}.

@@

\noindent where the subscript $j$ corrresponds to any neighboring pedestrian or 
the walls. The three forces $\mathbf{f}_d$, $\mathbf{f}_s$ and $\mathbf{f}_g$ 
are different in nature. The desire force $\mathbf{f}_d$ represents the 
acceleration (or deceleration) of the pedestrian due to his (her) own \textcolor{blue}{will}. 
The social force $\mathbf{f}_s$, instead, describes the tendency of the 
pedestrians to stay away from each other. The granular force $\mathbf{f}_g$ 
stands for \textcolor{red}{both} the sliding friction and the compression between 
pedestrians. \\

>>>>

The numerical \textcolor{blue}{parameters} for the ``faster is slower'' effect presented by 
Helbing and co-workers \textcolor{blue}{\sout{is somewhat cumbersome} appears to have some drawbacks} 
Ref.~\cite{helbing_2000,dorso_2017,dorso_2019}. Although a single 
parameter (say, the desired velocity $v_d$) is numerically varied to 
explain the phenomenon, the researcher \textcolor{blue}{\sout{looses sight} may loose sight} of the dimensionless 
``numbers'' that truly control the crowd dynamics. The \textcolor{blue}{set of parameters} in 
Ref.~\cite{helbing_2000} also misses the ``faster is faster'' effect reported 
to occur at very high pedestrian densities \cite{dorso_2017,haghani_2019}. 
\textcolor{blue}{Moreover}, the empirical fundamental diagram raises as a point of reference 
for the SFM control parameters \cite{helbing_2007,dorso_2017}. \\

The fundamental diagram exhibits the flux behavior for either low \textcolor{blue}{density} crowds 
(with \textcolor{blue}{rare} contacts between pedestrians) and highly dense crowds (dominated 
by \textcolor{blue}{two body interactions}). \textcolor{blue}{In the second case, crowds} experience a flux 
slowing down, but other behaviors are also possible \cite{helbing_2007,lohner_2018}. \textcolor{blue}{
\sout{In light of our previous hypothesis, we } We} may suspect that the  modeling of the 
``flux slowing down'' within the context of the SFM will require the proper 
setting of the (dimensionless) controlling parameters. We \textcolor{blue}{have already}
examined this working hypothesis in Ref.~\cite{dorso_2019}, but limiting the parameter exploration to 
the sliding friction, disregarding the body force. 
We now widen the investigation on the \textcolor{blue}{\sout{parameter settings} set of parameters} to include the one 
associated to the body force. We will  explore 
 the complex interplay between 
the body force and the sliding friction among pedestrians. Recall that the 
interplay dynamics is not directly controlled by the \textcolor{blue}{\sout{numerical setting} set of parameters}, but 
through dimensionless ``numbers'', where the model parameters appear mixed 
between each other. Thus, this step up offers a challenge to the 
``individualistic'' meaning of the parameter's \textcolor{blue}{\sout{setting} set}. \\

@@

The complex behavior of pedestrians features
 either his (her) feelings and 
the environmental conditions. The former is expressed, for example, by his 
(her) moving ``attitude'' (say,\textcolor{blue}{\sout{self-assuredness} relaxed, normal, aggressive, etc}).
The latter brings out the 
observed separation between pedestrians. Also, the ``contacts'' between 
individuals \textcolor{blue}{produce} some kind of ``unwanted'' slowing down. All these 
observed patterns are commonly quantified in the literature into a set of 
characteristic  parameters. \textcolor{red}{It is worth mentioning that the empirical
data reported in this Section corresponds to non-panicking situations in which pedestrians 
do not experience high anxiety levels (see below). }
The experimental meaning of these parameters is as follows.

\begin{itemize}
\item[(i)] The walking attitude of a pedestrian may appear somewhat 
  ``aggressive'' if he (she) reacts actively to unexpected behaviors 
\cite{lakoba_2005,helbing_1995}. The smaller the reaction time, the more 
 aggressive observed posture. The associated parameter to this 
behavior is the relaxation or characteristic time $\tau$ 
\cite{johansson_2009,helbing_2000}.

@@

\noindent where the subscript $j$ corrresponds to any neighboring pedestrian or 
the walls. The three forces $\mathbf{f}_d$, $\mathbf{f}_s$ and $\mathbf{f}_g$ 
are different in nature. The desire force $\mathbf{f}_d$ represents the 
acceleration (or deceleration) of the pedestrian due to his (her) own \textcolor{blue}{will}. 
The social force $\mathbf{f}_s$, instead, describes the tendency of the 
pedestrians to stay away from each other. The granular force $\mathbf{f}_g$ 
stands for \textcolor{blue}{both} the sliding friction and the compression between 
pedestrians. \\
